---
title: "test space"
author: "Brandon Wallace"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Ideas:

For filters:
1. by year in the date column 
2. by month in a calendar
3. by quarter

Three Tabs:
1. Employment
    - Unemployment rate over time
    - Share of working population
    - Median Household Income
2. Housing
    - Average Mortgage
    - National Home Price Index
    - Monthly Home Supply (numeric box)
3. GDP / Inflation
    - Inflation
    - Corporate Bond Yield (as table)
    - GDP per capita (numeric box)
    - Real GDP (numeric box)
    
or 

For numeric boxes and gauges
1. 3 - year average home price
2. 3 - year average home supply
3. 3 - year average household income
```{r}
library(readr)
macro_data <- read_csv("~/carnegie mellon - r shiny/hw2blwallac/hw2-blwallac/us-macro-data-2002-2022.csv")
```


```{r}
library(dplyr)
macro_data$DATE <- as.Date(macro_data$DATE, format = "%d/%m/%Y")
```


```{r}
summary(macro_data)
which(is.na(macro_data))
```


```{r}

macro_data$YEAR <- format(macro_data$DATE, "%Y")
macro_data$YEAR <- as.numeric(macro_data$YEAR)
macro_data$QUARTERS <- quarters(macro_data$DATE)  

```


```{r}
macro_data$`TOO HOT` <- ifelse(macro_data$`INFLATION(%)` < 2, 0, 1)
summary(macro_data)

```


```{r}
save(macro_data, file = "macro_data.rdata")
```


### Labor Market

Employment
    - Unemployment rate over time
    - Share of working population
    - Median Household Income
    
Filters - any date plus quarters 

```{r}

ggplot(data = macro_data, aes(x = `DATE`, y = `% SHARE OF WORKING POPULATION`)) +
  geom_smooth(color="red", size = 1.5) +
  labs(title = "Percent Share of Population in the Workforce", x = "Date", y = "(%) Share of Working Population")

```


```{r}

ggplot(data = macro_data, aes(x = `DATE`, y = `UNRATE(%)`)) +
  geom_line(color="black") + geom_point(color="blue") +
  scale_x_date(date_breaks = "years" , date_labels = "'%y") +
  labs(title = "Unemployment Rate Over Time", x = "Date", y = "Unemployment (%)")

```


### Housing Market 

2. Housing
    - National Home Price Index
    - Monthly Home Supply (numeric box)
    - Median Household Income (numeric box)
    - Household interest payment (numeric box)
    
Filters - slider dates 
    
```{r}

ggplot(data = macro_data, aes(x = `DATE`, y = `CSUSHPISA`)) +
  geom_line(color="black", linetype="dashed", size = 0.5) +
  scale_x_date(date_breaks = "years" , date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "U.S. National Home Price Index", x = "Date", y = "Home Price Index")

```


```{r}

mean(macro_data$`MED HOUSEHOLD INCOME`, na.rm=TRUE)
```


### GDP / Inflation
    - Inflation over time
    - read gdp and gdp per capita
    - Corporate Bond Yield (as table)

```{r}

ggplot(data = macro_data, aes(x = `DATE`, y = `INFLATION(%)`)) +
  geom_line(color="black", size = 0.5) +
  scale_x_date(date_breaks = "years" , date_labels = "%Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept= 2, color="orange", size=.5) +
  annotate(geom="text", x=as.Date("2008-5-29"), y=7, 
             label="U.S. Stock Market Collapse of 2008") +
    annotate(geom="point", x=as.Date("2008-7-29"), y=5.5, size=5, shape=21, fill="transparent") +
  labs(title = "Inflation Rate over Time", x = "Date", y = "Inflation (%)")
```


```{r}

ggplot(data = macro_data, aes(x = `DATE`)) +
  geom_line(aes(y = `GDP PER CAPITA`), color = "blue", size = 1.2) +
  geom_line(aes(y = `QUARTERLY REAL GDP`), color = "red", size = 1.2) + 
  scale_x_date(date_breaks = "years" , date_labels = "'%y") +
   annotate(geom="text", x=as.Date("2008-5-29"), y=60000, 
             label="GDP Per Capita") +
   annotate(geom="text", x=as.Date("2012-5-29"), y=21000, 
             label="Quarterly Real GDP") +
  labs(title = "Gross Domestic Product", x = "Date", y = "GDP")


```


```{r}

DT::datatable(data = macro_data,
                      options = list(pageLength = 50), 
                      rownames = FALSE) %>%
        formatStyle('TOO HOT', target = 'row', backgroundColor = styleEqual(c(0, 1), c('lightskyblue', 'orangered'))) 

```


```{r}
```


```{r}
```


```{r}
```

